>

### 2020-06-05 02_parks_data.Rmd
Repo cleanup
This includes a number of edits to the data elements related to switching the unit of analysis from tracts to block groups
Clean up file names and headers

---
title: "Parks Data"
author: "Greg Macfarlane"
date: "6/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(leaflet)
library(tidyverse)
library(RSocrata)
library(leaflet)
library(sf)
library(lubridate)
```

>>>>

---
title: "Parks Data"
---

```{r setup, cache = FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(leaflet)
library(tidyverse)
library(RSocrata)
library(leaflet)
library(sf)
library(lubridate)
```

### 2020-06-05 03_distances.Rmd

With `r nrow(open_space_points)` and `r nrow(bg_centroids)`, that suggests we
need to compute `r nrow(open_space_points) * nrow(bg_centroids)` shortest
paths. Even on a 20+ core process, this would take several weeks.  As a result,
we simplify the problem by first identifying the "closest" point of the park
by Euclidean distance and then computing the network path to that point.
This introduces some distortion where if the closest Euclidean point is across 
a river or interstate, the method calculates the network distance to *that*
point instead of the first point the person would encounter following the
network. This may end up overstating the true experienced distance.

>>>>

With `r nrow(park_points)` and `r nrow(tract_centroids)`, that suggests we
need to compute `r nrow(park_points) * nrow(bg_centroids)` shortest
paths. Even on a 20+ core process, this would take several weeks.  As a result,
we simplify the problem by first identifying the "closest" point of the park
by Euclidean distance and then computing the network path to that point.
This introduces some distortion where if the closest Euclidean point is across 
a river or interstate, the method calculates the network distance to *that*
point instead of the first point the person would encounter following the
network. This may end up overstating the true experienced distance.

### 2020-06-05 04_accessibility.Rmd

```{r load}
tracts <- read_rds("data/tracts.rds")
open_spaces <- read_rds("data/open_spaces.rds")
```

@@

read_dist_csv <- function(x){
  # warnings are from `inf` strings, which become converted to NA, which is okay.
  read_csv(x, col_types = "ccnn") %>% as_tibble()
}

blockgroup_distances <- lapply(blockgroup_files, read_dist_csv) %>% bind_rows()
tract_distances      <- lapply(tract_files,      read_dist_csv) %>% bind_rows()
```

@@

    # compute utility
    mutate(
      euc_access = -1.768 * log(euc_dist) + 0.373 * log(as.numeric(acres)),
      euc_multi  = euc_access + trails * 0.99 +  playgrounds * 0.26 + courts * 0.43,
      net_access = -1.768 * log(distance) + 0.373 * log(as.numeric(acres)),
      net_multi  = net_access + trails * 0.99 +  playgrounds * 0.26 + courts * 0.43
    ) %>%

>>>>

```{r load}
tracts <- read_rds("data/tracts.rds")
open_spaces <- read_rds("data/open_spaces.rds")
distances <- read_rds("data/distances.rds")
```

@@

read_dist_csv <- function(x){
  # warnings are from `inf` strings, which become converted to NA, which is okay.
  suppressWarnings(read_csv(x, col_types = "ccnn") %>% as_tibble())
}

blockgroup_distances <- lapply(blockgroup_files, read_dist_csv) %>% 
  bind_rows() %>% 
  left_join(distances$blockgroups %>% rename(euc_old = distance), 
            by = c("geoid", "park_id"))
tract_distances <- lapply(tract_files, read_dist_csv) %>% 
  bind_rows() %>% 
  left_join(distances$tracts %>% rename(euc_old = distance), 
            by = c("geoid", "park_id"))
```

@@

    # compute utility
    mutate(
      euc_access = -1.768 * log(euc_old) + 0.373 * log(as.numeric(acres)),
      euc_multi  = euc_access + trails * 0.99 +  playgrounds * 0.26 + courts * 0.43,
      net_access = -1.768 * log(distance) + 0.373 * log(as.numeric(acres)),
      net_multi  = net_access + trails * 0.99 +  playgrounds * 0.26 + courts * 0.43
    ) %>%

### 2020-06-05 manuscript/manuscript.Rmd

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there.  Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], but objective quantification 
may not be strictly necessary. For example, takano2002urban showed improved 
survival rates among senior citizens in Japan who self-reported having access
to qualitatively good walking spaces. Though encouraging, this result could
simply reveal self-selection: Those seniors who felt the walking space was good
were those who chose to enjoy it. Urban planners seeking to evaluate which
residents have good access to quality green space may desire objective and
therefore comparable measures.

@Dong2006 present a helpful mathematical heirarchy of
common quantitative mechanisms for calculating access, which we follow here.
Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_{ij}$ is 
within an "isochrone," or a geometric buffer defined by some threshold $D$,

@@

Another common access measure fits in this category, where distance-based
isochrones are replaced with political, geographical, or statistical boundaries.
These are especially common in comparative macroscopic studies, such as the
positive relationships found by @West2012 between density of park land in a
metropolitan area and self-reported population health indicators derived from
the Behavioral Risk Factor Surveillance System (BRFSS). On the other hand,
@Larson2016 found no relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities and self-reported scores
on the Gallup-Healthways Wellbeing Index. At the sub-metropolitan level, 
the @Stark2014 found a significant relationship in New York City between the percent
of a residential ZIP code that is park space and objectively measured body
mass index (BMI). It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of point $i$ is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_{ij})$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function and
$S_j = 1$.
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park influence walking more than
merely the distance to it. @hillsdon2006relationship found no relationship between
a gravity-based accessibility term and self-reported physical activity among middle-aged individuals in Norwich, England. @Zhang2011 developed a population-weighted,
gravity-based accessibility to parks metric with a national scope, though they
do not examine its correlation with health outcomes.

@@

As a travel impedance measure, we used the Euclidean distance between each tract's
population-weighted centroid and the nearest point on the edge of a park's border.
Euclidean distances have well-rehearsed limitations regarding their fidelity with
the underlying infrastructure network, etc. Network-based distances can also
suffer from challenges when applied to multimodal problems; these challenges are
exacerbated when the non-highway mode share is high, as is likely when considering
access to parks. We examined replacing the Euclidean distance with a network-based
distance; the choice of distance metric appeared inconsequential to our larger
findings in this case. We retained the Euclidean distance for computational simplicity. A
better metric of travel impedance may be a mode choice model logsum, which
weights all alternative travel alternatives against each other. We leave this as
a recommendation for future research.

>>>>

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there.  Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], but objective quantification 
may not be strictly necessary. For example, @takano2002urban showed improved 
survival rates among senior citizens in Japan who self-reported having access
to qualitatively good walking spaces. Though encouraging, this result could
simply reveal self-selection: Those seniors who felt the walking space was good
were those who chose to enjoy it. Urban planners seeking to evaluate which
residents have good access to quality green space may desire objective and
therefore comparable measures.

@Dong2006 present a helpful mathematical hierarchy of
common quantitative mechanisms for calculating access, which we follow here.
Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_{ij}$ is 
within an "isochrone," or a geometric buffer defined by some threshold $D$,

@@

Another common access measure fits in this category, where distance-based
isochrones are replaced with political, geographical, or statistical boundaries.
For example, @Mitchell2008 find that inequalities in health outcomes attributable to 
income are mitigated in statistical areas with a greater percentage of green space.
These are especially common in comparative macroscopic studies, such as the
positive relationships found by @West2012 between density of park land in a
metropolitan area and self-reported population health indicators derived from
the Behavioral Risk Factor Surveillance System (BRFSS). On the other hand,
@Larson2016 found no relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities and self-reported scores
on the Gallup-Healthways Wellbeing Index. At the sub-metropolitan level, 
@Stark2014 found a significant relationship in New York City between the percent
of a residential ZIP code that is park space and objectively measured body
mass index (BMI). It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the arbitrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of point $i$ is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_{ij})$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function and
$S_j = 1$.
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park influence walking more than
merely the distance to it. @hillsdon2006relationship found no relationship between
a gravity-based accessibility term and self-reported physical activity among middle-aged individuals in Norwich, England. @Zhang2011 developed a population-weighted,
gravity-based accessibility to parks metric with a national scope, though they
do not examine its correlation with health outcomes.

@@

As a travel impedance measure, we used the Euclidean distance between each tract's
population-weighted centroid and the nearest point on the edge of a park's border.
Euclidean distances have well-rehearsed limitations regarding their fidelity with
the underlying infrastructure network, etc. Network-based distances can also
suffer from challenges when applied to multimodal problems; these challenges are
exacerbated when the non-highway mode share is high, as is likely when considering
access to parks. We examined replacing the Euclidean distance with a network-based
distance; the choice of distance metric appeared inconsequential to our larger
findings in this case while substantially increasing the computational load required.
Additionally, we could not verify the accuracy of the underlying network data at 
the scale of this analysis. We therefore retain the Euclidean distance for
simplicity. A
better metric of travel impedance may be a mode choice model logsum, which
weights all alternative travel alternatives against each other. We leave this as
a recommendation for future research.
